---
title: "August 22"
output: html_document
---

```{r}
library(tidyverse)
```


19.5.5 Exercises

1. What does commas(letters, collapse = "-") do? Why? 
It doesn't work; it thinks we are putting collapse in as an actual argument as well as a "formal argument"
```{r}
commas <- function(...) stringr::str_c(..., collapse = ", ")
commas(letters, collapse = "-")
commas(letters[1:10])
```


It’d be nice if you could supply multiple characters to the pad argument, e.g. rule("Title", pad = "-+"). Why doesn’t this currently work? How could you fix it?

Our function assumes that pad has a length of 1. We can try to fix it by taking the length of pad into account

```{r}
rule <- function(..., pad = "-+~") {
  title <- paste0(...)
  width <- getOption("width") - nchar(title) - 5 
  cat(title, " ", stringr::str_dup(pad, width/nchar(pad)), "\n", sep = "")
}
rule("Important output")
```

What does the trim argument to mean() do? When might you use it?

It is a fraction of observations to be trimmed from each side of x before the mean is computed. You can use it if you have a lot of outliers on each side

The default value for the method argument to cor() is c("pearson", "kendall", "spearman"). What does that mean? What value is used by default?
They are different types of coefs. Default is pearson


# 20.3.5 Exercises

1. Describe the difference between is.finite(x) and !is.infinite(x).
is.finite considers any number that is not Inf to be finite. is.finite will consider NA to be not finite.
is.infinite considers only Inf and -Inf to be infinite. But then, !is.infinite will consider NA to not be infinite.

2. Read the source code for dplyr::near() (Hint: to see the source code, drop the ()). How does it work?
```{r}
dplyr::near
```

It allows a tolerance level for how different the numbers can be

3. A logical vector can take 3 possible values. How many possible values can an integer vector take? How many possible values can a double take? Use google to do some research.

?Not infinite?

4. Brainstorm at least four functions that allow you to convert a double to an integer. How do they differ? Be precise.



5. What functions from the readr package allow you to turn a string into logical, integer, and double vector?
parse_logical, parse_integer, and parse_number
