---
title: "August 30"
output: html_document
---

Aug 30: rest of Chapter 20

20.4.5 Subsetting

[] is the subsetting function for vectors and is called like x[a]

[[ only ever extracts a single element, and always drops names. It’s a good idea to use it whenever you want to make it clear that you’re extracting a single item, as in a for loop. 

20.4.6 Exercises

1. What does mean(is.na(x)) tell you about a vector x? What about sum(!is.finite(x))?
mean(is.na(x)) tells you what proportion of NAs there are in x. 
sum(!is.infinite(x)) tells you how many objects there are that are not -Inf or Inf

2. Carefully read the documentation of is.vector(). What does it actually test for? Why does is.atomic() not agree with the definition of atomic vectors above?
is.vector() checks if x has any attributes besides name. Maybe to check for more complex things?
is.atomic checks if x is one of the 6 atomic types, but it's ok to have other attributes

3. Compare and contrast setNames() with purrr::set_names().
purrr::set_names has additional checks to make sure input is valid, like must be a vector and must be same length

4. Create functions that take a vector as input and returns:
The last value. Should you use [ or [[?
```{r}
x[[length(x)]]
```
The elements at even numbered positions.
```{r}
x[0:(length(x) - 1) %% 2 == 0]
```

Every element except the last value.
```{r}
x[-length(x)]
```

Only even numbers (and no missing values).
```{r}
x[!is.na(x) && x %% 2 == 0]
```


5. Why is x[-which(x > 0)] not the same as x[x <= 0]?
edge case: NA. which(NA > 0) is false, so opposite of that is true.
But x <= NA is still false.

6. What happens when you subset with a positive integer that’s bigger than the length of the vector? What happens when you subset with a name that doesn’t exist?
```{r}
x <- 0:5
x[9]
x[["cool"]]
```
# 20.5 Recursive vectors (lists)

Subsetting a list

[] extracts a sublist. The result will always be a list
[[]] extracts a single component and removes one level of hierarchy from the list
$ extracts named elements of the list. Equivalent to [[""]]

What happens if you subset a tibble as if you’re subsetting a list? What are the key differences between a list and a tibble?
# borrowed answer
Subsetting a tibble works the same way as a list; a data frame can be thought of as a list of columns. The key different between a list and a tibble is that a tibble (data frame) has the restriction that all its elements (columns) must have the same length.

# 20.7.4 Exercises

1. What does hms::hms(3600) return? How does it print? What primitive type is the augmented vector built on top of? What attributes does it use?
```{r}
x <- hms::hms(3600)
typeof(x)
attributes(x)
class(x)
```


2. Try and make a tibble that has columns with different lengths. What happens?

```{r}
tibble(1:5, 1)
tibble(1:5, 1:3)
```

It works if the larger column is a multiple of the smaller one

3. Based on the definition above, is it ok to have a list as a column of a tibble?
I think so
