---
title: "May 17"
output: 
  html_document: 
    keep_md: yes
---

5.6.7 Exercises

```{r}
library(nycflights13)
library(tidyverse)

not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

2. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).
```{r 5.6.7 2}
not_cancelled %>% group_by(dest) %>% summarise(n())
not_cancelled %>% group_by(tailnum) %>% summarise(distance = sum(distance))
```

3. Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?
```{r} 
flights %>% summarise(sum(is.na(dep_delay)))
flights %>% summarise(sum(is.na(arr_delay)))
```
There are more NAs in arr_delay than dep_delay. Why? I don't know
It seems like dep_delay is more important because a cancelled flight will never take off...r ight?

4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

```{r}
cancelled <- flights %>% mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(year, month, day) %>% summarise(cancelled_percent = mean(cancelled),
                                           avg_delay = mean(dep_delay, na.rm=TRUE))
cancelled

ggplot(cancelled, aes(avg_delay, cancelled_percent)) + geom_point()
```


5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))
```{r}
flights %>% group_by(carrier) %>% summarise(delay = mean(dep_delay, na.rm = TRUE)) %>% arrange(desc(delay))
#F9 is the worst carrier

# Incomplete Challenge: 
# Number of flights to each destination by each carrier
(by_dest <- flights %>% group_by(carrier, dest) %>% summarise(n()))

flights %>% group_by(carrier, dest) %>% summarise(delay=mean(dep_delay, na.rm=TRUE)) %>% summarise(mean(delay, na.rm=TRUE))
```

6. What does the sort argument to count() do. When might you use it?
```{r}
not_cancelled %>% count(dest, sort=T)
```
It sorts by count. Seems like it could have been useful earlier when we used arrange

# 5.6.1 Exercises
1. The operations will apply to groups, not the whole data frame
2. Which plane (tailnum) has the worst on-time record?
```{r}
flights %>% group_by(tailnum) %>% summarise(mean_delay = mean(arr_delay, na.rm=TRUE)) %>% arrange(desc(mean_delay))
#I hate flying on N844MH
```
3. What time of day should you fly if you want to avoid delays as much as possible? 
```{r}
flights %>% group_by(hour) %>% summarise(mean_delay = mean(arr_delay, na.rm=TRUE)) %>% arrange(mean_delay)
#7am wakin up in the morning
```
4. For each destination, compute the total minutes of delay. For each flight compute the proportion of the total delay for its destination. 
```{r}
not_cancelled %>% group_by(dest) %>% mutate(total_delay = sum(arr_delay), delay_percent = arr_delay / total_delay)
```

5. Use lag() to explore how the delay of a flight is related to the delay of the immediately preceeding flight.
```{r}
not_cancelled %>% mutate(lag_delay = lag(dep_delay)) %>%
  ggplot(aes(dep_delay, lag_delay)) + geom_point()
```
6. Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  mutate(median = median(air_time), speed=(air_time - median), normalized=speed/median) %>% 
  arrange(normalized) %>%
  select(dest, median, normalized, air_time)

# Doesn't this not take into account origin? ..isn't that important?
```


7. Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.
rank the carriers? what?

```{r}
#Counts rows, or number of flights per carrier per destination
flights %>% group_by(dest, carrier) %>% count(carrier)


#Counts rows, or number of destinations per carrier
flights %>% group_by(dest, carrier) %>% count(carrier) %>% group_by(carrier) %>% count()

```

8. For each plane, count the number of flights before the first delay of greater than 1 hour.
```{r}
flights %>% group_by(tailnum) %>% mutate(delayed = dep_delay > 60) %>% mutate(beforeflights = cumsum(delayed)) %>% filter(beforeflights < 1) %>% count()
```

