---
title: "September 13"
output: html_document
---

```{r}
library(tidyverse)
library(stringr)
```

map() makes a list.
map_lgl() makes a logical vector.
map_int() makes an integer vector.
map_dbl() makes a double vector.
map_chr() makes a character vector.

# 21.5.3 Exercises

Write code that uses one of the map functions to:
Compute the mean of every column in mtcars.
```{r}
mtcars %>%
  map_dbl(mean)
```

Determine the type of each column in nycflights13::flights.
```{r}
map(nycflights13::flights, class)
```

Compute the number of unique values in each column of iris.
```{r}
map_int(iris, ~length(unique(.)))
```

Generate 10 random normals for each of  
```{r}
map(c(-10, 0, 10, 100), rnorm, n = 10)
```

How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?
```{r}
map_lgl(iris, is.factor)
```

What happens when you use the map functions on vectors that aren’t lists? What does map(1:5, runif) do? Why?
```{r}
map(1:5, runif)
```
It doesn't have to be a list, it will still map functions to vectors


What does map(-2:2, rnorm, n = 5) do? Why? What does map_dbl(-2:2, rnorm, n = 5) do? Why?
```{r}
map(-2:2, rnorm, n=5)
map_dbl(-2:2, rnorm, n=5)
```
map(...) takes 5 samples from the normal distribution of means -2, -1, 0, 1, 2
map_dbl() throws an error because it expects rnorm to return a length 1 numeric vector. But rnorm returns 5 because it samples 5 times.

Rewrite map(x, function(df) lm(mpg ~ wt, data = df)) to eliminate the anonymous function.
```{r}
map(x, ~ lm(mpg ~ wt, data=df))
```

# 21.9.3 Exercises

Implement your own version of every() using a for loop. Compare it with purrr::every(). What does purrr’s version do that your version doesn’t?
```{r}
every2() <- function(.data, .predicate, ...){
  for (i in .data) {
    if (!.predicate){
      return(FALSE)
    }
  }
  return(TRUE)
}
  
```
purrr::every : "A single predicate function, a formula describing such a predicate function, or a logical vector of the same length as .x. Alternatively, if the elements of .x are themselves lists of objects, a string indicating the name of a logical element in the inner lists. Only those elements where .p evaluates to TRUE will be modified."

So, it can also accept a logical vector (like one you'd make from map_lgl)

Create an enhanced col_sum() that applies a summary function to every numeric column in a data frame.

```{r}
col_sum2 <- function(data, f, ...) {
  data2 <- keep(data, is.numeric)
  map(data2, f, ...)
}
```

A possible base R equivalent of col_sum() is:
```{r}
col_sum3 <- function(df, f) {
  is_num <- sapply(df, is.numeric)
  df_num <- df[, is_num]

  sapply(df_num, f)
}
```
But it has a number of bugs as illustrated with the following inputs:
```{r}
df <- tibble(
  x = 1:3, 
  y = 3:1,
  z = c("a", "b", "c")
)
# OK
col_sum3(df, mean)
# Has problems: don't always return numeric vector
col_sum3(df[1:2], mean)
col_sum3(df[1], mean)
col_sum3(df[0], mean)
```

What causes the bugs?

```{r}
class(col_sum3(df, mean))
class(col_sum3(df[1:2], mean))
class(col_sum3(df[1], mean))
class(col_sum3(df[0], mean))
```
It tries to return a list if the input is a blank vector? Not sure what the other issues are
