---
title: "September 6"
output: html_document
---

Sep 6: 21.1 - 21.4

```{r}
library(tidyverse)
library(nycflights13)
library(stringr)
```

# 21.2.1 Exercises

1. Write for loops to:

Compute the mean of every column in mtcars.
```{r}
output <- vector("double", ncol(mtcars))
for (i in seq_along(mtcars)) {        
  output[[i]] <- mean(mtcars[[i]])
}
output
```

Determine the type of each column in nycflights13::flights.
```{r}
output <- vector("list", ncol(nycflights13::flights))
for (i in seq_along(flights)) {        
  output[[i]] <- class(flights[[i]])
}
output
```

Compute the number of unique values in each column of iris.
```{r}
output <- vector("integer", ncol(iris))
names(output) <- names(iris)
for (i in seq_along(iris)) {        
  output[[i]] <- length(unique(iris[[i]]))
}
output
```

Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors:
```{r}
out <- ""
for (x in letters) {
  out <- stringr::str_c(out, x)
}

stringr::str_c(letters, collapse = "")

x <- sample(100)
sd <- 0
for (i in seq_along(x)) {
  sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))

sd(x)

x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}
```

Combine your function writing and for loop skills:

Write a for loop that prints() the lyrics to the children’s song “Alice the camel”.
```{r}
for(i in 5:0) {
  if(i != 1) {
    print(str_c("Alice the camel has ", i, " humps."))
  } else {
    print("Alice the camel has 1 hump.")
  }
  if( i !=0) {
    print("So go, Alice, go.")
  } else {
    print("Now Alice is a horse.")
  }
}
```

# 21.3.5 Exercises

1. Imagine you have a directory full of CSV files that you want to read in. You have their paths in a vector, files <- dir("data/", pattern = "\\.csv$", full.names = TRUE), and now want to read each one with read_csv(). Write the for loop that will load them into a single data frame.
```{r}
output <- vector("list", length(files))
for(i in seq_along(files)) {
  output[[i]] <- read_csv(files[[i]])
}
output <- bind_rows(output)
```

2. What happens if you use for (nm in names(x)) and x has no names? What if only some of the elements are named? What if the names are not unique?
```{r}
x <- 1:3
print(names(x))

for (nm in names(x)) {
  print(nm)
  print(x[[nm]])
}
```

If no names, the loop doesn't run. 
If an element is unnamed, the name is actually ""
For duplicate names, it will return only the first element


Write a function that prints the mean of each numeric column in a data frame, along with its name. For example, show_mean(iris) would print:

```{r}
show_mean <- function(df) {
  for(i in names(df)) {
    if(is.numeric(df[[i]])) {
      cat(str_c(i, ": ", mean(df[[i]]), "\n"))
    }
  }
}

show_mean(iris)
```

What does this code do? How does it work?
```{r}
trans <- list( 
  disp = function(x) x * 0.0163871,
  am = function(x) {
    factor(x, labels = c("auto", "manual"))
  }
)
for (var in names(trans)) {
  mtcars[[var]] <- trans[[var]](mtcars[[var]])
}
```

It changes the $disp and $am columns. It multiplies disp by 0.016 and factorizes am to auto or manual 


# 21.4.1 Exercises

Read the documentation for apply(). In the 2d case, what two for loops does it generalise?
It applies to rows or columns

Adapt col_summary() so that it only applies to numeric columns You might want to start with an is_numeric() function that returns a logical vector that has a TRUE corresponding to each numeric column.
```{r}
col_summary <- function(df, fun) {
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    if(is.numeric(df[[i]])) {
      out[i] <- fun(df[[i]])
    }
  }
  out
}
```

