---
title: "August 9"
output: html_document
---

```{r}
library(tidyverse)
library(forcats)
library(lubridate)
library(nycflights13)
```


15.3.1 Exercise

Explore the distribution of rincome (reported income). What makes the default bar chart hard to understand? How could you improve the plot?
```{r}
gss_cat %>% ggplot(aes(rincome)) + geom_bar() + coord_flip()
```

What is the most common relig in this survey? What’s the most common partyid?
```{r}
gss_cat %>%
  count(relig) %>%
  arrange(-n)
```

Which relig does denom (denomination) apply to? How can you find out with a table? How can you find out with a visualisation?
```{r}
levels(gss_cat$denom)

gss_cat %>%
  count(relig, denom) %>%
  ggplot(aes(x = relig, y = denom, size = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90))
```
```{r}
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )


ggplot(relig_summary, aes(tvhours, fct_reorder(relig, tvhours))) +
  geom_point()
```

15.5.1 Exercises

How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?
```{r}
gss_cat %>% 
  mutate(partyid = 
           fct_collapse(partyid,
                        other = c("No answer", "Don't know", "Other party"),
                        rep = c("Strong republican", "Not str republican"),
                        ind = c("Ind,near rep", "Independent", "Ind,near dem"),
                        dem = c("Not str democrat", "Strong democrat"))) %>%
  count(year, partyid)  %>%
  group_by(year) %>%
  mutate(p = n / sum(n)) %>%
  ggplot(aes(x = year, y = p,
             colour = fct_reorder2(partyid, year, p))) +
  geom_point() +
  geom_line() +
  labs(colour = "Party ID.")
```

How could you collapse rincome into a small set of categories?
```{r}
library("stringr")
gss_cat %>%
  mutate(rincome = 
           fct_collapse(
             rincome,
             `Unknown` = c("No answer", "Don't know", "Refused", "Not applicable"),
             `Lt $5000` = c("Lt $1000", str_c("$", c("1000", "3000", "4000"),
                                              " to ", c("2999", "3999", "4999"))),
             `$5000 to 10000` = str_c("$", c("5000", "6000", "7000", "8000"),
                                      " to ", c("5999", "6999", "7999", "9999"))
           ))
```
