---
title: "R-club-Apr-26"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 3.2.4

```{r 3.2.4 Exercises, echo = FALSE}
ggplot(data = mpg)
paste("Rows: ", nrow(mtcars))
paste("Columns: ", ncol(mtcars))
print("drv describes the 'drive'")
ggplot(mpg) + geom_point(aes(hwy, cyl))
ggplot(mpg) + geom_point(aes(class, drv))
```


1. The color part should have been outside of the aes wrapper.
2. Categorical variables: manufacturer, model, trans, drv, fl, class
      Continuous variables: displ, year, cyl, cty, hwy
      Generally, continuous variables will have a class of double / int, and categorical ones will be characters
3. Continuous variables can't be mapped to shape. For color or size, they will be mapped to a gradient depending on their value. Categorical variables will be ordered alphabetically and mapped to the same gradients, but will probably be meaningless/arbitrary. For shape, the first 6 variables will be assigned shapes. 
4. Both aesthetics will be applied.
5. Stroke is like a paint stroke, sort of density visualizer
6. It will map to the output of whatever you put

1. It will make a subplot for every possible value.
2. The empty cells mean there were no data points with that combination of the two faceted variables
3. '. ~ var' means do not facet by row. 'var ~ .' means do not facet by column.
4. Advantages: Possibly easier to see trends for an individual subgroup. Color might be better if you have a large dataset and want to compare everything side by side.
5. Facet wrap controls the visual output of the panels on your screen. Facet grid does not have nrow/ncol because those are determined by the variables that you are faceting by.
6. Our screens are usually in 'landscape' mode, so there is more room to display more columns.
