---
title: "July 19"
output: html_document
---

Finish chapter 14 and embedded exercises

```{r}
library(tidyverse)
library(stringr)
```

To determine if a character vector matches a pattern, use `str_detect()`
A common use of str_subset() is to select the elements that match the pattern. 

For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.

Find all words that start or end with x.

```{r}
str_subset(words, "(^x)|(x$)")
```


Find all words that start with a vowel and end with a consonant.

```{r}
str_subset(words, "^[aeiou].*[^aeiou]$")
```

Are there any words that contain at least one of each different vowel?
```{r}
str_subset(words, "a") %>% str_subset("e") %>% str_subset("i") %>% str_subset("o") %>% str_subset("u")
```

What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)

```{r}
df <- tibble(
  word = words, 
)
df <- df %>%
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]"),
    proportion = vowels / consonants
  ) 

arrange(df, desc(vowels)) %>% head()
arrange(df, desc(proportion)) %>% head()
```

14.4.4.1 Exercises

1. Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.
```{r}
number <- "(one|two|three|four) ([^ ]+)"
has_number <- sentences %>% str_subset(number)

has_number %>% str_match(number)
```

Find all contractions. Separate out the pieces before and after the apostrophe.

```{r}
contraction <- "([^ ]+)\'([^ ]+)"
sentences %>% str_subset(contraction) %>% str_match(contraction)
```

14.4.5 Replacing Matches

Replace all forward slashes in a string with backslashes. 
## I can't figure this out. 

```{r}

hello <- "hello//"

hello %>% str_replace_all("/", "\\\\")
```

Implement a simple version of str_to_lower() using replace_all().
## This one either. Maybe by subtracting from uppercase to get to the lowercase version

```{r}
str_replace(sentences, "(A-Z)",tolower(("\\1"))) %>% head()
```


Switch the first and last letters in words. Which of those strings are still words?
```{r}
switched <- words %>% str_replace("(^.)(.*)(.$)", "\\3\\2\\1")
switchedcol <- str_c("^", switched, "$",collapse = "|")

words %>% str_subset(switchedcol)

```

14.4.6.1 Exercises

Split up a string like "apples, pears, and bananas" into individual components.
```{r}
fruits <- "apples, pears, and bananas"
fruits %>% str_split("(, and )|(, )")
```

Why is it better to split up by boundary("word") than " "?
```{r}
gruits <- "I like to eat apples, pears, and bananas!"
gruits %>% str_split(" ")
gruits %>% str_split(boundary("word"))
```
Using boundary("word") will not capture the punctuation that is attached to words


What does splitting with an empty string ("") do? Experiment, and then read the documentation.
```{r}
gruits %>% str_split("")
```

It splits into every single character. It is equivalent to splitting for boundary("character")

14.5.1 Exercises

How would you find all strings containing \ with regex() vs. with fixed()? 
#? Still don't understand fixed()
```{r}
str_subset(c("a\\b", "ab"), regex("\\\\"))
str_subset(c("a\\b", "ab"), fixed("\\"))
```


What are the five most common words in sentences?
```{r}
sentences %>% str_extract_all(boundary("word", ignore_case=TRUE)) %>%
  unlist() %>% #Because it gives many lists of lists 
  str_to_lower() %>%
  tibble() %>%
  set_names("word") %>%
  group_by(word) %>%
  count(sort= TRUE) %>%
  head(5)
```


Find the stringi functions that:

```{r}
apropos("stri")
apropos("duplicate")
apropos("stri_rand")
```

Count the number of words. `stri_count()`
Find duplicated strings. `stri_duplicated()`
Generate random text. "stri_rand_lipsum"  "stri_rand_shuffle" "stri_rand_strings"
How do you control the language that stri_sort() uses for sorting? use locale=
