---
title: "June 14"
output: 
  html_document: 
    keep_md: yes
---

```{r, include=FALSE}
library(tidyverse)
```

## 12.2.1 
1. Each variable should have its own column. So instead of having one column for 1994, and one column for 1995, we should have one column for year. This may require "duplicating" rows, but this allows observations to all be in one row. 

2. 
```{r}
table1 %>% 
  mutate(rate = cases / population * 10000)

cases <- table2[table2$type=="cases",]$count
country <- table2[table2$type=="cases",]$country
year <- table2[table2$type=="cases",]$year
population <- table2[table2$type=="population",]$count
table2_rate <- tibble(country = country,
       year = year,
       rate = cases / population)
table2_rate


#table4a$1999 won't work because 1999 begins with a number. Instead use [[]]
table4_rate <- tibble(country = table4a[["country"]],
       `1999` = table4a[["1999"]] / table4b[["1999"]],
       `2000` = table4b[["2000"]] / table4b[["2000"]])
```
3. 
```{r}
table2 %>% 
  filter(type == "cases") %>%
  ggplot(aes(year, count)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```

## 12.3.3
1. 
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`)
```
Unsure what this question is. The order is changed, maybe because half comes before year alphabetically?

2. 
```{r}
table4a %>% 
  gather( `1999`, `2000`, key = "year", value = "cases")
#> Error in combine_vars(vars, ind_list): Position must be between 0 and n
```
Numbers in gather default to column numbers, not names

3. 
```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```
Phillip's age shows up twice. Maybe we can add a rep column, to say this is the second time we measured Phillip's height, at a different age. 


4. 
```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
preg %>% gather(key = sex, value = count, male, female)
```
We need to gather because each row represents two observations.
Variables: Pregnancy, Sex, Count

## 12.4.3 Exercises
1. 
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra="drop")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
```
extra tells you what to do with excess values. Fill is the opposite. 

2. Remove lets you remove the old variable you just separated, similar to transmute. If you wanted to keep both, like in mutate, you would do remove=false. 

3. Unite puts everything together. You can't unite by position, separator, or group... but you can separate in different ways. 
